Do moral communities have a spatial dimension? A spatial exploratory analysis of places of worship and violent crime in the city of Recife, Brazil

### 2019-10-15 Moral_Communities_and_Crime_v1.Rmd
Completed first draft of analysis/results and conclusions.

Yet another fruitful avenue for research, and one that has only recently begun to be explored, is the presence of environmental attributes that can help to reinforce moral norms, such as schools and churches [e.g., @Abdullah2018amenities; @Davignon2015christian; @Furr2010metric; @Traunmuller2011moral]. It is thus that in a recent paper, Warner and Konkel [-@Warner2019neighborhood] note that the role of places of worship, as distinct entities from the members of the congregations, have received less attention in empirical and theoretical research for their deterrence potential. The role of these institutions might be particularly important in places where formal state institutions lack the means or the will to enforce norms - as is the case in Brazil.

@@

For the analysis, we use Principal Component Analysis, a data reduction technique, to obtain an indicator of Socio-Economic Deprivation. The loadings of the first factor, which accounts for almost 54% of the variance, are shown in Table \ref{tab:sed-descriptive-statistics}. The loadings of this factor indicate that high Socio-Economic Deprivation is a combination of (in terms of importance): high levels of poverty, high levels of unemployment, and low median income, followed by high levels of family disruption, in particular proportion of families led by young single mothers.

@@

kable(sed_descriptives,
      "latex",
      booktabs = TRUE,
      digits = 2,
      escape = FALSE,
      caption = "\\label{tab:sed-descriptive-statistics}Descriptive statistics of some key socio-economic and demographic variables in Recife",
      col.names = linebreak(c("Variable", 
                              "Min", 
                              "2nd Quartile", 
                              "Median", 
                              "Mean", 
                              "3rd Quartile", 
                              "Max", "PC Factor 1\n Loadings")),
      row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "scaled_down")) %>%
  footnote(general = c("The variable for young single mothers is for women aged 15-25",
                       "The first principal component accounts for 53.91% of the variance"))

@@

Figure \ref{fig:plot-sed-as-quintiles} displays the geography of Socio-Economic Deprivation in Recife, after classifying _setores_ by quintiles, whereby the "High" class corresponds to _setores_ in the top 20% of the Socio-Economic Deprivation indicator, and the "Low" class corresponds to _setores_ in the bottom 20% of the Socio-Economic Deprivation indicator. The figure shows a veritable mosaic of affluence and deprivation, with high deprivation areas directly in contact, and in some cases even completely surrounded, by low deprivation areas. This geographical pattern of unequality, on the other hand, seems to be characteristic of Brazilian metropolitan regions, where enclaves of wealth and _favelas_ (i.e., urban slums) can be found in close proximity [see for example @Feitosa2007global].

@@

Calculate the intensity of other ppp by quadrat:
```{r quadrat-analysis, include=FALSE}

# Calculate the intensity of the various point patterns by quadrat:

@@

@@

@@

@@

@@

@@

>>>>

Yet another fruitful avenue for research, and one that has only recently begun to be explored, is the presence of environmental attributes that can help to reinforce moral norms, such as schools and churches [e.g., @Abdullah2018amenities; @Davignon2015christian; @Furr2010metric; @Traunmuller2011moral]. It is thus that in a recent paper, Warner and Konkel [-@Warner2019neighborhood] note that the role of places of worship, as distinct entities from the members of the congregations, have received less attention in empirical and theoretical research for their deterrence potential. The role of these institutions might be particularly important in places where formal state institutions lack the means or the will to enforce norms in a consistent way - as is the case in Brazil [@Garmany2014space].

@@

For the analysis, we use Principal Component Analysis, a data reduction technique, to obtain an indicator of Socio-Economic Deprivation, essentially the first principal component in the output. The loadings of this factor (Table \ref{tab:sed-descriptive-statistics}) indicate that high Socio-Economic Deprivation is a combination of (in terms of importance): high levels of poverty, high levels of unemployment, and low median income, followed by high levels of family disruption, in particular proportion of families led by young single mothers. This factor accounts for almost 54% of the variance..

@@

kable(sed_descriptives,
      "latex",
      booktabs = TRUE,
      digits = 2,
      escape = FALSE,
      align = c("l", "c", "c", "c", "c", "c", "c", "c"),
      caption = "\\label{tab:sed-descriptive-statistics}Descriptive statistics of some key socio-economic and demographic variables in Recife",
      col.names = linebreak(c("Variable", 
                              "Min", 
                              "2nd Quartile", 
                              "Median", 
                              "Mean", 
                              "3rd Quartile", 
                              "Max", 
                              "PC Factor 1\n Loadings")),
      row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "scaled_down")) %>%
  footnote(general = c("The variable for young single mothers is for women aged 15-25",
                       "The first principal component accounts for 53.91% of the variance"))

@@

Figure \ref{fig:plot-sed-as-quintiles} displays the geography of Socio-Economic Deprivation in Recife, after classifying _setores_ by quintiles, whereby the "High" class corresponds to _setores_ in the top 20% of the Socio-Economic Deprivation indicator, and the "Low" class corresponds to _setores_ in the bottom 20% of the Socio-Economic Deprivation indicator. The figure shows a veritable mosaic of affluence and deprivation, with high deprivation areas directly in contact, and in some cases almost completely surrounded, by low deprivation areas. This geographical pattern of unequality, on the other hand, seems to be characteristic of Brazilian metropolitan regions, where enclaves of wealth and _favelas_ (i.e., urban slums) can be found in close proximity [see for example @Feitosa2007global].

@@

Once the indicator of Socio-Economic Deprivation has been obtained, as outlined above, it can be used for conditional quadrat analysis. As previously described, the quadrats are irregular areas that are defined based on a covariate, in this case the quintiles of the Socio-Economic Deprivation indicator. In essence, this entails counting the number of events at each level of the covariate, and then calculating an estimator of the intensity.

```{r quadrat-analysis, include=FALSE}

# Calculate the intensity of the various point patterns by quadrat, conditional on Socio-Economic Deprivation:

@@

Figure \ref{fig:plot-crime-quadrat} shows the results of conditional quadrat counts for Lethal and Intentional Violent Crime. The dotted line indicates the global intensity, simply the total number of criminal events divided by the area of the region. It can be seen that the intensity of crime clearly increases with Socio-Economic Deprivation, and the intensity in areas with the highest levels of Socio-Economic Deprivation is `r crime5_intensity[5]/crime5_intensity[1] * 100` % higher than in places with the lowest levels of Socio-Economic Deprivation. This is as expected, as research has consistently found correlations between Socio-Economic Deprivation and a variety of negative health outcomes, including homicide [@Ichihara2018area]. A $\chi^2$ test of independence on the quadrat counts yields a value of $p\leq 0.0001$ (four degrees of freedom), which comfortably rejects the null hypothesis of homogeneity of intensity of crime by level of Socio-Economic Deprivation.

@@

Next, we repeat the analysis of conditional quadrat counts, but now for the six types of events of reference. As discussed above, this included three types of places of worship which are of interest from the perspective of moral communities, and three types of commercial establishments that we posit are morally neutral. The results of estimating the intensity of the process by means of conditional quadrat counts are shown in Figure \ref{fig:plot-events-quadrat}.

The first thing that we note are the variations in global intensity of places of worship, with Evangelical places of worship being the most intense of the three. There are no clear variations in the locational pattern of Catholic places of worship by level of Socio-Economic Deprivation. Evangelical places of worship, on the other hand, tend to be found more frequently in places with middle-low and middle levels of Socio-Economic Deprivation, a pattern also discernible, albeit barely, for Spiritist churches. Interestingly (see Table \ref{tab:quadrat-test-statistics}) $\chi^2$ tests of independence on the conditional quadrat counts fail to reject the hypothesis of homogeneity for Catholic and Spiritist churches, and it is only in the case of Evangelical places of worship that we detect the possibility of locational patterns that vary by level of Socio-Economic Deprivation.

Three types of commercial establishments also appear to display inhomogeneous locational patterns (Table \ref{tab:quadrat-test-statistics}), with the $\chi^2$ test of independence emphaticaly rejecting the null hypothesis. As seen in Figure \ref{fig:plot-events-quadrat}, ice cream shops tend to be found more frequently in places with middle-low Socio-Economic Deprivation, and less frequently in places with middle-high Socio-Economic Deprivation. Pharmacies show a clearer trend, with locational patterns that tend to favor places with low Socio-Economic Deprivation. Finally, supermarkets are found less frequently in places with middle-high and high Socio-Economic Deprivation.

@@

```{r table-quadrat-test-statistics, echo=FALSE}

quad.tests <- data.frame(Event = c("Catholic",
                                   "Evangelical",
                                   "Spiritist",
                                   "Ice Crem Shop",
                                   "Pharmacy",
                                   "Supermarket"),
                         rbind(quadrat.test(catholic5), 
                               quadrat.test(evangelic5), 
                               quadrat.test(spiritist5),
                               quadrat.test(ice_cream5),
                               quadrat.test(pharmacies5),
                               quadrat.test(super_markets5))) %>%
  transmute(Event,
            Statistic = as.numeric(statistic),
            degrees_of_freedom = as.numeric(parameter),
            p_value = as.numeric(p.value)) %>%
  mutate(p_value = ifelse(p_value >= 0.001, paste0(round(p_value, 4)), "$<0.001$"))

kable(quad.tests,
      "latex",
      booktabs = TRUE,
      digits = 3,
      escape = FALSE,
      align = c("l", "c", "c", "c"),
      caption = "\\label{tab:quadrat-test-statistics}$\\chi^2$ tests of independence of conditional quadrat counts",
      col.names = linebreak(c("Event", 
                              "Statistic", 
                              "Degrees of Freedom", 
                              "p-value"))) %>%
  kable_styling(bootstrap_options = c("striped", "scaled_down"))
```

@@

The preceding analysis provides a valuable backdrop. As seen there, there is a clear distribution of criminal events, increasing on par with the level of Socio-Economic deprivation. Of the three different classes of places of worship, only Evangelical churches display a locational pattern that is commensurate, with Socio-Economic Deprivation, albeit the pattern is distinct from criminal events, with Evangelical churches found less frequently in both high and low Socio-Economic Deprivation areas.

The analysis does not answer the question, yet, of possible covariations between crime, on the one hand, and places of worship and commercial establishments, on the other. In this subsection we implement the relative distribution estimator of intensity of Lethal and Intentional Violent Crime. The covariate $Z(u)$ in this analysis is distance to one of a place of worship or a commercial establishment:
$$
\lambda_k(s)=\rho(Z_k(s))
$$
with:
$$
Z_k(s) = d_{sk}
$$
In the expression above, $d_{s_k}$ is distance from location $s$ to an event of class $k$, and $k=\{\text{Catholic, Evangelical, Spiritist, Ice Cream Shop, Pharmacy, Supermarket}\}$.

@@

```{r relative-distribution, include=FALSE, cache=TRUE}

# The estimates of intensity are calculated with `rhohat`:

rho_catholic <- rhohat(events_recife_list$`Violent Crime`, z_catholic)
rho_evangelic <- rhohat(events_recife_list$`Violent Crime`, z_evangelical)
rho_spiritist <- rhohat(events_recife_list$`Violent Crime`, z_spiritist)
rho_ice_cream <- rhohat(events_recife_list$`Violent Crime`, z_ice_cream)
rho_pharmacies <- rhohat(events_recife_list$`Violent Crime`, z_pharmacies)
rho_supermarkets <- rhohat(events_recife_list$`Violent Crime`, z_super_markets)
```

```{r prepare-relative-distribution-data, include=FALSE}

# Rename the variables in the results to collect in a single dataframe:

rho_catholic <- rho_catholic %>% as.data.frame() %>% dplyr::rename(z = z_catholic)
rho_evangelic <- rho_evangelic %>% as.data.frame() %>% rename(z = z_evangelica)
rho_spiritist <- rho_spiritist %>% as.data.frame() %>% dplyr::rename(z = z_spiritist)
rho_ice_cream <- rho_ice_cream %>% as.data.frame() %>% rename(z = z_ice_cream)
rho_pharmacies <- rho_pharmacies %>% as.data.frame() %>% rename(z = z_pharmacies)
rho_supermarkets <- rho_supermarkets %>% as.data.frame() %>% rename(z = z_super_mark)

# Collect into dataframe for convenience of plotting:

rho_all <- rbind(data.frame(rho_catholic, Type = "Catholic", Class = "Place of Worship"), 
                 data.frame(rho_evangelic, Type = "Evangelical", Class = "Place of Worship"),
                 data.frame(rho_spiritist, Type = "Spiritist", Class = "Place of Worship"),
                 data.frame(rho_ice_cream, Type = "Ice Cream Shops", Class = "Business"), 
                 data.frame(rho_pharmacies, Type = "Pharmacies", Class = "Business"), 
                 data.frame(rho_supermarkets, Type = "Supermarkets", Class = "Business")) 
```

```{r analysis-relative-distribution, include=FALSE}

# Extract some useful information from the relative distribution functions

# First, find the max intensity for Evangelical:

max_int_evangelical <- max(filter(rho_all, Type == "Evangelical")$rho)

# Next, the distance at which the intensity of crime with respect to distance to Evangelical is max:

dist_max_int_evangelical <- filter(rho_all, Type == "Evangelical")[filter(rho_all, Type == "Evangelical")$rho == max_int_evangelical,"z"]

# Next, find the approximate intensity at that distance for the other types of events. First, what is the nearest distance to the distance at which the intensity with respect to Evangelical is max?

near_dist_catholic_i <- which.min(abs(filter(rho_all, Type == "Catholic")$z - dist_max_int_evangelical))

# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row. Use the two rows to obtain a linear approximation of the intensity at the desired distance

if(filter(rho_all, Type == "Catholic")[near_dist_catholic_i, "z"] > dist_max_int_evangelical){
  interv <- filter(rho_all, Type == "Catholic")[(near_dist_catholic_i-1):near_dist_catholic_i,] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  catholic_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
}else{
  interv <- filter(rho_all, Type == "Catholic")[near_dist_catholic_i:(near_dist_catholic_i + 1),] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  catholic_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
}

# Repeat for Spiritist

near_dist_spiritist_i <- which.min(abs(filter(rho_all, Type == "Spiritist")$z - dist_max_int_evangelical))

# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row. Use the two rows to obtain a linear approximation of the intensity at the desired distance

if(filter(rho_all, Type == "Spiritist")[near_dist_spiritist_i, "z"] > dist_max_int_evangelical){
  interv <- filter(rho_all, Type == "Spiritist")[(near_dist_spiritist_i-1):near_dist_spiritist_i,] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  spiritist_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
}else{
  interv <- filter(rho_all, Type == "Spiritist")[near_dist_spiritist_i:(near_dist_spiritist_i + 1),] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  spiritist_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
}

# Repeat for ice cream shop

near_dist_ice_cream_i <- which.min(abs(filter(rho_all, Type == "Ice Cream Shops")$z - dist_max_int_evangelical))

# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row. Use the two rows to obtain a linear approximation of the intensity at the desired distance

if(filter(rho_all, Type == "Ice Cream Shops")[near_dist_ice_cream_i, "z"] > dist_max_int_evangelical){
  interv <- filter(rho_all, Type == "Ice Cream Shops")[(near_dist_ice_cream_i-1):near_dist_ice_cream_i,] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  ice_cream_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
}else{
  interv <- filter(rho_all, Type == "Ice Cream Shops")[near_dist_ice_cream_i:(near_dist_ice_cream_i + 1),] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  ice_cream_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
}

# Repeat for pharmacy

near_dist_pharmacies_i <- which.min(abs(filter(rho_all, Type == "Pharmacies")$z - dist_max_int_evangelical))

# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row. Use the two rows to obtain a linear approximation of the intensity at the desired distance

if(filter(rho_all, Type == "Pharmacies")[near_dist_pharmacies_i, "z"] > dist_max_int_evangelical){
  interv <- filter(rho_all, Type == "Pharmacies")[(near_dist_pharmacies_i - 1):near_dist_pharmacies_i,] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  pharmacies_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
}else{
  interv <- filter(rho_all, Type == "Pharmacies")[near_dist_pharmacies_i:(near_dist_pharmacies_i + 1),] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  pharmacies_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
}

# Repeat for supermarkets

near_dist_supermarkets_i <- which.min(abs(filter(rho_all, Type == "Supermarkets")$z - dist_max_int_evangelical))

# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row. Use the two rows to obtain a linear approximation of the intensity at the desired distance

if(filter(rho_all, Type == "Supermarkets")[near_dist_supermarkets_i, "z"] > dist_max_int_evangelical){
  interv <- filter(rho_all, Type == "Supermarkets")[(near_dist_supermarkets_i - 1):near_dist_supermarkets_i,] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  supermarkets_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
}else{
  interv <- filter(rho_all, Type == "Supermarkets")[near_dist_supermarkets_i:(near_dist_supermarkets_i + 1),] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  supermarkets_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
}
```

The results of this analysis are shown in Figure \ref{fig:plot-relative-distribution}. The global estimator for the intensity of crime is shown in this figure as a dotted line. Three types of places of worship are shown in solid lines, and three types of commercial establishments are shown in dashed lines. Each relative distribution function is shown with its corresponding 95% confidence bands.

Several interesting things emerge from inspection of this figure. First, we notice that with the exception of Spiritist churches (which has an intensity significantly lower than the global intensity at a distance of zero), the intensity of crime at for other churches and commercial establishments is close, but higher than, the global intensity of crime. In general, the intensity tends to increase at short distances; however, this effect is much more marked for Evangelical places of worship. The intensity of crime in the proximity of these places of worship grows rapidly within a short distance, reaching a peak of `r signif(max_int_evangelical, 5)` at a distance of `r round(dist_max_int_evangelical, 2)` m. At that distance, the intensity of crime with respect to Evangelical places of worship is `r round(max_int_evangelical/catholic_int_at_evangelical_max_int * 100, 2)`% higher than the intensity of Catholic places of worship; `r round(max_int_evangelical/spiritist_int_at_evangelical_max_int * 100, 2)`% higher than the intensity of crime with respect to Spiritist churches; `r round(max_int_evangelical/ice_cream_int_at_evangelical_max_int * 100, 2)`% higher than the intensity with respect to ice cream shops; `r round(max_int_evangelical/pharmacies_int_at_evangelical_max_int * 100, 2)`% higher than the intensity with respect to pharmacies; and `r round(max_int_evangelical/supermarkets_int_at_evangelical_max_int * 100, 2)`% higher than the intensity with respect to supermarkets.

Intriguingly, after reaching its peak intensity, the intensity of crime with respect to Evangelical places of worship declines sharply, until at distances of approximately 200 m it is lower than for other places of worship and commercial establishments, and at distances of approximately 270 m it is lower than the global intensity of crime. This does not happen with other places of worship or commercial establishments: the intensity of crime with respect to these locations remains higher than the global intensity of crime for the interval of distances examined. Another interesting observation is that the intensity of crime with respect to Catholic places of worship is not substantially different when compared to the intensity of crime with respect to any of the commercial establishments considered.

These results, while intriguing, beg the question of whether variations in Socio-Economic Deprivation may have a confounding effect on the relative distributions; for instance, if as seen in the case of conditional quadrat counts, there is some overlap in the locational patterns of, say, Evangelical places of worship and crime with respect to Socio-Economic Deprivation. Next section explores this question.

```{r figure-plot-relative-distribution, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:plot-relative-distribution}Intensity of crime as a function of distance to a selected place of worship or commercial establishment; bands are 95% confidence interval and the dotted line indicates the global intensity of crime"}
# Plot all estimates. The dotted line is the intensity if the intensity was constant (this is a summary measure of the `crime.ppp` object). Select only distance less than 500 m:

ggplot(data = filter(rho_all, z <= 500)) +
  geom_ribbon(aes(x = z, ymin = lo, ymax = hi, fill = Type), alpha = 0.2) + 
  geom_line(aes(x = z, y = rho, color = Type, linetype = Class), size = 1) +
  geom_abline(intercept = 7.575416e-06, slope = 0, linetype = "dotted") +
  #xlim(c(0, 500)) +
  theme_tufte()
```

@@

In this section, the preceding analysis regarding the relative distribution of crime using distance to places of worship and commercial establishments is repeated, after introducing a baseline function. The interpretation of $\rho(Z(s))$ in this case is as follows: instead of the raw intensity at the value of $z$, it is the intensity _relative to the intensity according to the level of Socio-Economic Deprivation_. Therefore, $rho(z)=1$ matches the intensity according to Socio-Economic Deprivation; values of $rho(z)>1$ indicate a higher intensity than explained by Socio-Economic Deprivation; and values of $rho(z)<1$ correspond to lower intensities than explained by Socio-Economic Deprivation.

```{r relative-distribution-with-baseline, include=FALSE}

#The estimates of intensity are calculated with `rhohat`, but now with a baseline intensity function that depends on median income:

rho_catholic_bl <- rhohat(events_recife_list$`Violent Crime`, z_catholic, baseline = sed.img)
rho_evangelic_bl <- rhohat(events_recife_list$`Violent Crime`, z_evangelical, baseline = sed.img)
rho_spiritist_bl <- rhohat(events_recife_list$`Violent Crime`, z_spiritist, baseline = sed.img)
rho_ice_cream_bl <- rhohat(events_recife_list$`Violent Crime`, z_ice_cream, baseline = sed.img)
rho_pharmacies_bl <- rhohat(events_recife_list$`Violent Crime`, z_pharmacies, baseline = sed.img)
rho_supermarkets_bl <- rhohat(events_recife_list$`Violent Crime`, z_super_markets, baseline = sed.img)
```

```{r prepare-relative-distribution-with-baseline-data, include=FALSE}

# Rename variables to combine dataframes:

rho_catholic_bl <- rho_catholic_bl %>% 
  as.data.frame() %>% 
  dplyr::rename(z = z_catholic)
rho_evangelic_bl <- rho_evangelic_bl %>% 
  as.data.frame() %>% 
  rename(z = z_evangelica)
rho_spiritist_bl <- rho_spiritist_bl %>% 
  as.data.frame() %>% 
  dplyr::rename(z = z_spiritist)
rho_ice_cream_bl <- rho_ice_cream_bl %>% 
  as.data.frame() %>% 
  rename(z = z_ice_cream)
rho_pharmacies_bl <- rho_pharmacies_bl %>% 
  as.data.frame() %>% 
  rename(z = z_pharmacies)
rho_supermarkets_bl <- rho_supermarkets_bl %>% 
  as.data.frame() %>% 
  rename(z = z_super_mark)

# Collect into dataframe for convenience of plotting:

rho_all_bl <- rbind(data.frame(rho_catholic_bl, Type = "Catholic", Class = "Place of Worship"), 
                    data.frame(rho_evangelic_bl, Type = "Evangelical", Class = "Place of Worship"),
                    data.frame(rho_spiritist_bl, Type = "Spiritist", Class = "Place of Worship"),  
                    data.frame(rho_ice_cream_bl, Type = "Ice Cream Shops", Class = "Business"), 
                    data.frame(rho_pharmacies_bl, Type = "Pharmacies", Class = "Business"), 
                    data.frame(rho_supermarkets_bl, Type = "Supermarkets", Class = "Business")) 
```

```{r analysis-relative-distribution-with-baseline, include=FALSE}

# Extract some useful information from the relative distribution functions

# First, find the max intensity for each type of event:

max_int_evangelical_bl <- max(filter(rho_all_bl, Type == "Evangelical")$rho)
max_int_catholic_bl <- max(filter(rho_all_bl, Type == "Catholic")$rho)
max_int_spiritist_bl <- max(filter(rho_all_bl, Type == "Spiritist")$rho)
max_int_ice_cream_bl <- max(filter(rho_all_bl, Type == "Ice Cream Shops")$rho)
max_int_pharmacy_bl <- max(filter(rho_all_bl, Type == "Pharmacies")$rho)
max_int_supermarket_bl <- max(filter(rho_all_bl, Type == "Supermarkets")$rho)

# Next, the distance at which the intensity of crime with respect to distance to Evangelical is max:

dist_max_int_evangelical_bl <- filter(rho_all_bl, Type == "Evangelical")[filter(rho_all_bl, Type == "Evangelical")$rho == max_int_evangelical_bl,"z"]

# Next, find the approximate intensity at that distance for the other types of events. First, what is the nearest distance to the distance at which the intensity with respect to Evangelical is max?

near_dist_catholic_i <- which.min(abs(filter(rho_all_bl, Type == "Catholic")$z - dist_max_int_evangelical_bl))

# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row. Use the two rows to obtain a linear approximation of the intensity at the desired distance

if(filter(rho_all_bl, Type == "Catholic")[near_dist_catholic_i, "z"] > dist_max_int_evangelical_bl){
  interv <- filter(rho_all_bl, Type == "Catholic")[(near_dist_catholic_i-1):near_dist_catholic_i,] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  catholic_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
}else{
  interv <- filter(rho_all_bl, Type == "Catholic")[near_dist_catholic_i:(near_dist_catholic_i + 1),] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  catholic_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
}

# Repeat for Spiritist

near_dist_spiritist_i <- which.min(abs(filter(rho_all_bl, Type == "Spiritist")$z - dist_max_int_evangelical_bl))

# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row. Use the two rows to obtain a linear approximation of the intensity at the desired distance

if(filter(rho_all_bl, Type == "Spiritist")[near_dist_spiritist_i, "z"] > dist_max_int_evangelical_bl){
  interv <- filter(rho_all_bl, Type == "Spiritist")[(near_dist_spiritist_i-1):near_dist_spiritist_i,] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  spiritist_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
}else{
  interv <- filter(rho_all_bl, Type == "Spiritist")[near_dist_spiritist_i:(near_dist_spiritist_i + 1),] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  spiritist_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
}

# Repeat for ice cream shop

near_dist_ice_cream_i <- which.min(abs(filter(rho_all_bl, Type == "Ice Cream Shops")$z - dist_max_int_evangelical_bl))

# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row. Use the two rows to obtain a linear approximation of the intensity at the desired distance

if(filter(rho_all_bl, Type == "Ice Cream Shops")[near_dist_ice_cream_i, "z"] > dist_max_int_evangelical_bl){
  interv <- filter(rho_all_bl, Type == "Ice Cream Shops")[(near_dist_ice_cream_i-1):near_dist_ice_cream_i,] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  ice_cream_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
}else{
  interv <- filter(rho_all_bl, Type == "Ice Cream Shops")[near_dist_ice_cream_i:(near_dist_ice_cream_i + 1),] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  ice_cream_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
}

# Repeat for pharmacy

near_dist_pharmacies_i <- which.min(abs(filter(rho_all_bl, Type == "Pharmacies")$z - dist_max_int_evangelical_bl))

# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row. Use the two rows to obtain a linear approximation of the intensity at the desired distance

if(filter(rho_all_bl, Type == "Pharmacies")[near_dist_pharmacies_i, "z"] > dist_max_int_evangelical_bl){
  interv <- filter(rho_all_bl, Type == "Pharmacies")[(near_dist_pharmacies_i - 1):near_dist_pharmacies_i,] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  pharmacies_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
}else{
  interv <- filter(rho_all_bl, Type == "Pharmacies")[near_dist_pharmacies_i:(near_dist_pharmacies_i + 1),] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  pharmacies_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
}

# Repeat for supermarkets

near_dist_supermarkets_i <- which.min(abs(filter(rho_all_bl, Type == "Supermarkets")$z - dist_max_int_evangelical_bl))

# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row. Use the two rows to obtain a linear approximation of the intensity at the desired distance

if(filter(rho_all_bl, Type == "Supermarkets")[near_dist_supermarkets_i, "z"] > dist_max_int_evangelical_bl){
  interv <- filter(rho_all_bl, Type == "Supermarkets")[(near_dist_supermarkets_i - 1):near_dist_supermarkets_i,] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  supermarkets_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
}else{
  interv <- filter(rho_all_bl, Type == "Supermarkets")[near_dist_supermarkets_i:(near_dist_supermarkets_i + 1),] %>%
    dplyr::select(z, rho)
  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
  supermarkets_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
}
```

The results of the analysis are shown in Figure \ref{fig:plot-relative-distribution-with-baseline}. The general trends shown by the relative distribution functions are similar to those seen above, before the introduction of a baseline function. We notice that at short distances the intensity of crime as a function of distance to the various places of worship and commercial establishments is higher than the baseline intesity (i.e., the intensity according to Socio-Ecomomic Deprivation). The only exception is intensity as a function of distance to a Spiritist church, which at short distances is not significantly different from the baseline intensity (see the confidence bands).

The peak intensity as a function of distance to Evangelical churches is `r round(max_int_evangelical_bl, 2)` times the baseline intensity; for Catholic places of worship, the highest intesity is `r round(max_int_catholic_bl, 2)` times the baseline intensity; and the intensity reaches a peak of `r round(max_int_spiritist_bl, 2)` times the baseline intensity when the covariate is distance to Spiritist churches. Turning now to inspect the three types of commercial establishments, we see that the highest intensities are `r round(max_int_ice_cream_bl, 2)`, `r round(max_int_pharmacy_bl, 2)`, and `r round(max_int_supermarket_bl, 2)` times the intensity according to Socio-Economic Deprivation for ice cream shops, pharmacies, and supermarkets, respectively.
